---
title: "Part 2: Web scraping practice"
author: Dan Turner (dturner@u.northwestern.edu)
---

```{r run this first}
library(rvest) # Web scraping
library(tidyverse) # Data wrangling
library(RCurl) # Download files from the internet

```

# Part 2: **Web scraping practice**

## Challenge 1
Modify the rule below to list the titles of all the blog posts on the first page found at the URL:

```{r Challenge 1 answer}

url <- "https://www.r-bloggers.com/"

#rule <- "body > div > div.wrapper-corporate > div > div > article:nth-child(1) > header > h3"

rule <- "body > div > div.wrapper-corporate > div > div > article > header > h3" # answer

read_html(url) %>% 
  html_nodes(rule) %>%
  html_text()

```


## Challenge 2
Modify the rule below to make a dataframe consisting of the titles, links, author, and date. The author and date will require you to use Inspector view to build and test two more rules.

```{r Challenge 2 answer}

url <- "https://www.r-bloggers.com/"

title_rule <- "body > div > div.wrapper-corporate > div > div > article > header > h3 > a"

author_rule <- "body > div > div.wrapper-corporate > div > div > article > header > p > span > a"

date_rule <- "body > div > div.wrapper-corporate > div > div > article > header > p"


titles <- read_html(url) %>% 
  html_nodes(title_rule) %>%
  html_text()

links <- read_html(url) %>% 
  html_nodes(title_rule) %>%
  html_attr('href')

authors <- read_html(url) %>% 
  html_nodes(author_rule) %>%
  html_text()

dates <- read_html(url) %>% 
  html_nodes(date_rule) %>%
  html_text()

df <- data.frame(titles, links, authors, dates, stringsAsFactors = FALSE)

View(df)

```



## Challenge 3
Now that we can extract data from one page, let's make sure we can get every page.

Write a function that lists every page of blog posts.

```{r Challenge 3 sample answer}

# let's find out how many pages there are
page_rule <- "body > div > div.wrapper-corporate > div > div > a:nth-child(25)"

# the url of the last page has its page number in it
page.count <- read_html(url) %>% 
  html_nodes(page_rule) %>%
  html_text()

# make it numeric
page.count <- as.numeric(gsub(",","",page.count))


# I think this is the most simple solution
all_the_links <- paste0("https://www.r-bloggers.com/page/", 1:page.count, "/")

tail(all_the_links)
```

If you want to scrape all_the_links, you can now identify the wanted information, write a rule to extract it, then loop all_the_links to build the dataset.



